### 2017-06-09 sheets/talk-6/sheet-6.tex
First part of sample solution 6.10

\begin{answer}
  \begin{itemize}
    \item[``$\Rightarrow$'']
      \emph{Counterexample:}
      Let $\cat{A} := \Set, E := \emptyset, X := \nat, Y := \mathbb{B}$.
      Define $f(n) := \text{even}(n)$ and $g(n) := \text{odd}(n)$. $i$ is the (empty) inclusion function.
      Then
      \[ \xymatrix{
        E \ar[r]^i \ar[d]_i &
        X \ar[d]^g \\
        X \ar[r]_f &
        Y
      } \]
      is clearly a pullback because no number is even and odd at the same time.
      Now let $A := \nat$, $f_1(n) := n$ and $f_2(n) := n+1$.
      Then the diagram
      \[ \xymatrix{
        A \ar[r]_{f_2} \ar[d]^{f_1} & Y \ar[d]_{g} \\
        X \ar[r]^f & Z
      } \]
      commutes because $\text{even}(n) \Leftrightarrow \text{odd}(n+1)$.
      But there exists no function $\bar{f} \from \nat \to \emptyset$ at all.
    \item[``$\Leftarrow$'']
      \begin{proof}
        Let
        \[ \xymatrix{
          E \ar[r]^i \ar[d]_i &
          X \ar[d]^g \\
          X \ar[r]_f &
          Y
        } \]
        be the pullback of the diagram
        $ \xymatrix{
          X \ar[r]^f & Y & X\ar[l]_g \\
        } $.
        We show that $\xymatrix@1{E \ar[r]^i &X \ar@<.5ex>[r]^{f} \ar@<-.5ex>[r]_g &Y}$ is the equalizer of $\xymatrix@1{X \ar@<.5ex>[r]^{f} \ar@<-.5ex>[r]_g &Y}$.
        First we note that the diagramm $\xymatrix@1{E \ar[r]^i &X \ar@<.5ex>[r]^{f} \ar@<-.5ex>[r]_g &Y}$ is indeed a fork as it commutes by assumption.
        Let
        $ \xymatrix{
          A \ar[r]^f & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
        } $
        be a fork. Then
        \[ \xymatrix{
          A \ar[r]_{f} \ar[d]^{f} & X \ar[d]_{t} \\
          X \ar[r]^s & Y
        } \]
        commutes. Therefore exists a unique map $\bar{f}\from A \to P$ such that
        \[ \xymatrix{
          A \ar[drr]^{f} \ar[ddr]_{f} \ar@{.>}[dr]|{\bar{f}} & & \\
          & P \ar[r]_{i} \ar[d]^{i} & X \ar[d]_{t} \\
          & X \ar[r]^s & Y
        } \]
        commutes. We show that this $\bar{f}$ is the only map such that
        \[ \xymatrix{
          A \ar[rd]^f \ar@{.>}[dd]|{\bar{f}} & & \\
          & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \\
          P \ar[ru]_i & &
        } \]
        commutes. First it is clear that this diagramm commutes for $\bar{f}$. Now let $\tilde{f}$ be another map such that
        \[ \xymatrix{
          A \ar[rd]^f \ar@{.>}[dd]|{\tilde{f}} & & \\
          & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \\
          P \ar[ru]_i & &
        } \]
        commutes. But then
        \[ \xymatrix{
          A \ar[drr]^{f} \ar[ddr]_{f} \ar@{.>}[dr]|{\tilde{f}} & & \\
          & P \ar[r]_{i} \ar[d]^{i} & X \ar[d]_{t} \\
          & X \ar[r]^s & Y
        } \]
        also commutes. By the assumption that $\bar{f}$ is unique we have $\bar{f} = \tilde{f}$. \qedhere
      \end{proof}
  \end{itemize}
\end{answer}

@@

>>>>

\begin{answer}
  \begin{itemize}
    \item[``$\Rightarrow$'']
      \emph{Counterexample:}
      Let $\cat{A} := \Set, E := \emptyset, X := \nat, Y := \mathbb{B}$.
      Define $f(n) := \text{even}(n)$ and $g(n) := \text{odd}(n)$. $i$ is the (empty) inclusion function.
      Then
      \[ \xymatrix{
        E \ar[r]^i \ar[d]_i &
        X \ar[d]^g \\
        X \ar[r]_f &
        Y
      } \]
      is clearly a pullback because no number is even and odd at the same time.
      Now let $A := \nat$, $f_1(n) := n$ and $f_2(n) := n+1$.
      Then the diagram
      \[ \xymatrix{
        A \ar[r]_{f_2} \ar[d]^{f_1} & Y \ar[d]_{g} \\
        X \ar[r]^f & Z
      } \]
      commutes because $\text{even}(n) \Leftrightarrow \text{odd}(n+1)$.
      But there exists no function $\bar{f} \from \nat \to \emptyset$ at all.
    \item[``$\Leftarrow$'']
      \begin{proof}
        Let
        \[ \xymatrix{
          E \ar[r]^i \ar[d]_i &
          X \ar[d]^g \\
          X \ar[r]_f &
          Y
        } \]
        be the pullback of the diagram
        $ \xymatrix{
          X \ar[r]^f & Y & X\ar[l]_g \\
        } $.
        We show that $\xymatrix@1{E \ar[r]^i &X \ar@<.5ex>[r]^{f} \ar@<-.5ex>[r]_g &Y}$ is the equalizer of $\xymatrix@1{X \ar@<.5ex>[r]^{f} \ar@<-.5ex>[r]_g &Y}$.
        First we note that the diagram $\xymatrix@1{E \ar[r]^i &X \ar@<.5ex>[r]^{f} \ar@<-.5ex>[r]_g &Y}$ is indeed a fork as it commutes by assumption.
        Let
        $ \xymatrix{
          A \ar[r]^f & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y
        } $
        be a fork. Then
        \[ \xymatrix{
          A \ar[r]_{f} \ar[d]^{f} & X \ar[d]_{t} \\
          X \ar[r]^s & Y
        } \]
        commutes. Therefore exists a unique map $\bar{f}\from A \to P$ such that
        \[ \xymatrix{
          A \ar[drr]^{f} \ar[ddr]_{f} \ar@{.>}[dr]|{\bar{f}} & & \\
          & P \ar[r]_{i} \ar[d]^{i} & X \ar[d]_{t} \\
          & X \ar[r]^s & Y
        } \]
        commutes. We show that this $\bar{f}$ is the only map such that
        \[ \xymatrix{
          A \ar[rd]^f \ar@{.>}[dd]|{\bar{f}} & & \\
          & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \\
          P \ar[ru]_i & &
        } \]
        commutes. First it is clear that this diagram commutes for $\bar{f}$. Now let $\tilde{f}$ be another map such that
        \[ \xymatrix{
          A \ar[rd]^f \ar@{.>}[dd]|{\tilde{f}} & & \\
          & X \ar@<.5ex>[r]^s \ar@<-.5ex>[r]_t & Y \\
          P \ar[ru]_i & &
        } \]
        commutes. But then
        \[ \xymatrix{
          A \ar[drr]^{f} \ar[ddr]_{f} \ar@{.>}[dr]|{\tilde{f}} & & \\
          & P \ar[r]_{i} \ar[d]^{i} & X \ar[d]_{t} \\
          & X \ar[r]^s & Y
        } \]
        also commutes. By the assumption that $\bar{f}$ is unique we have $\bar{f} = \tilde{f}$. \qedhere
      \end{proof}
  \end{itemize}
\end{answer}

@@

\begin{answer}
  \begin{proof}
    We give names to the arrows and objects:
    \[ \xymatrix{
      A \ar[r]^h \ar[d]^j & B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
      D \ar[r]^m & E \ar[r]^n & F
    } \]
    Now let the pullback square
    \[ \xymatrix{
      B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
      E \ar[r]^n & F
    } \]
    be the pullback of $\xymatrix{E \ar[r]^n & F & C \ar[l]_l}$.
    We show both directions of the equivalence:
    \begin{itemize}
      \item[``$\Rightarrow$'']
        Let the pullback square
        \[ \xymatrix{
          A \ar[r]^{i \of h} \ar[d]^j & C \ar[d]^l \\
          D \ar[r]^{n \of m} & F
        } \]
        be the pullback of $\xymatrix{D \ar[r]^{n \of m} & F & C \ar[l]_l}$.
        Let
        \[ \xymatrix{
          Z \ar[r]^{f_1} \ar[d]^{f_2} & B \ar[d]^k \\
          D \ar[r]^{m} & E
        } \]
        be another pullback square. Now
        \[ \xymatrix{
          Z \ar[r]^{f_1} \ar[d]^{f_2} & B \ar[r]^{i} \ar[d]^k & C \ar[d]^l \\
          D \ar[r]^{m} & E \ar[r]^n & F
        } \]
        commutes. Now because the outer square is a pullback there is a unique map $\bar{f} \from Z \to A$, such that
        \[ \xymatrix{
          Z \ar@/^/[drrr]^{i \of f_1}
            \ar@{.>}[dr]|{\bar{f}}
            \ar@/_/[ddr]_{f_2} & & \\
            & A \ar[r]_h \ar[d]^j & B \ar[r]_{i} & C \ar[d]_l \\
            & D \ar[r]^{m} & E \ar[r]^n & F
        } \]
        commutes. We show that this $\bar{f}$ is the unique map such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{f_1}
            \ar@{.>}[dr]|{\bar{f}}
            \ar@/_/[ddr]_{f_2} \\
          & A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
          & D \ar[r]^m & E
        } \]
        commutes. We have to show $j \of \bar{f} = f_2$ and $h \of \bar{f} = f_1$. The first equatation holds by assumption.

        Now we define two maps
        \begin{alignat*}{4}
          g_1 &\from Z \to C&  \qquad g_1 &:= i \of f_1 \\
          g_2 &\from Z \to E&         g_2 &:= m \of j \of \bar{f}
        \end{alignat*}
        and show that the diagram
        \[ \xymatrix{
          Z \ar[r]^{g_1} \ar[d]^{g_2} & C \ar[d]^l \\
          E \ar[r]^n & F
        } \]
        commutes and is therefore a pullback square of the right-hand square:
        \begin{alignat*}{2}
          l \of g_1
          &= l \of i \of f_1 \\
          &= l \of i \of h \of \bar{f} \\
          &= n \of m \of j \of \bar{f} \\
          &= n \of g_2
        \end{alignat*}
        Now because the right-hand square is a pullback there is a unique map $\bar{g} \from Z \to B$, such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{g_1}
            \ar@{.>}[dr]|{\bar{g}}
            \ar@/_/[ddr]_{g_2} \\
          & B \ar[r]^i \ar[d]^k & C \ar[d]^l \\
          & E \ar[r]^n & F
        } \]
        commutes.
        By showing that this diagram comutes for $\bar{g} = f_1$ and for $\bar{g} = h \of \bar{f}$ we show $h \of \bar{f} = f_2$.
        \begin{itemize}
          \item[1.]We have $i \of f_1 = g_1$ by definition and $k \of f_1 = m \of f_2 = m \of j \of \bar{f} = g_2$.
          \item[2.]We have $(i \of h) \of \bar{f} = i \of f_2$ and $(k \of h) \of \bar{f} = (m \of j) \of \bar{f}$.
        \end{itemize}

        We have shown that the diagram
        \[ \xymatrix{
          Z \ar@/^/[drr]^{f_1}
            \ar@{.>}[dr]|{\bar{f}}
            \ar@/_/[ddr]_{f_2} \\
          & A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
          & D \ar[r]^m & E
        } \]
        commutes for the given $\bar{f}$. We now have to prove that this $\bar{f}$ is unique.
        We assume a map $\tilde{f} \from Z \to A$ such that
        \[ \xymatrix{
          Z \ar@/^/[drr]^{f_1}
            \ar@{.>}[dr]|{\tilde{f}}
            \ar@/_/[ddr]_{f_2} \\
          & A \ar[r]^h \ar[d]^j & B \ar[d]^k \\
          & D \ar[r]^m & E
        } \]
        commutes. But than we have another map such that the diagram for the outer square commutes:
        \[ \xymatrix{
          Z \ar@/^/[drrr]^{i \of f_1}
            \ar@{.>}[dr]|{\tilde{f}}
            \ar@/_/[ddr]_{f_2} & & \\
            & A \ar[r]_h \ar[d]^j & B \ar[r]_{i} & C \ar[d]_l \\
            & D \ar[r]^{m} & E \ar[r]^n & F
        } \]
        We conclude that $\bar{f} = \tilde{f}$.
      \item[``$\Leftarrow$'']
    \end{itemize}
  \end{proof}
\end{answer}
